>

### 2019-10-01 manuscript/manuscript.Rmd
Rewrite literature review to focus more on accessibility methods.

Metropolitan-level analyses such as those described above do not capture the
within-city variation in accessibility to parks that exist in many cities. Some
cities with large amounts of total greenspace may nevertheless unequally
distribute this space throughout the city, leading to areas with poor access.
Conversely, a city with a smaller overall proportion of greenspace may give all
of its citizens better access to sufficient greenspace to meet their needs or
wants for physical activity. Considering access at the neighborhood level within
a single city may also eliminate some regional or cultural fixed effects
affecting metropolitan-level analyses. In a study of neighborhoods in New York
City, @Stark2014 found that a higher proportion of park space in a neighborhood
was associated with a lower BMI for its residents, and lower park cleanliness
scores were associated with a higher BMI, suggesting that increasing the supply
of clean, well-maintained parks could positively impact adult BMI.

@@

Loosely defined, the *accessibility* of an arbitrary place describes the ease
with which people can accomplish activities there. Accessibility is an abstract
concept with tempting quantifiability, perhaps explaining the proliferation of
calculation mechanisms. @Dong2006 present a helpful mathematical heirarchy of
common mechanisms, which we follow here. Consider a person residing at
point $i$ in a city with parks $j \in 1 \ldots J$. An analyst might consider
point $i$ as "having access" to a park if the distance to the park $d_ij$  is less than some threshold
$D$,

\begin{equation}\ref{eq:isochrone}
A_i = \begin{cases} 
      1 & d_{ij}\leq D\\
      0 & d_{ij} > D
   \end{cases}
\end{equation}

Variations of this isochrone-based framework are easily derived and relatively common. 
@ParkScore is a sophisticated application of this approach where $d_{ij}$ is a
carefully calculated 10-minute network-bound walk, with a demographic analysis of
areas within and without this threshold. In other applications $D$ might be more variably defined,
such as "Does a park exist within the census tract?". The indicator function could also be
modified into the total number of parks or percent of park space within radius $D$ of point $i$.

A somewhat more holistic approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of a point is the denominator of the 
gravity formulation of a trip distribution model,

\begin{equation}\ref{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}

where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_ij)$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function.

A third accessibility mechanism is a utility-based specification, termed as such
for being derived from random utility choice theory. Consider that an individual
living at point $i$ is choosing a park for a recreation activity. If we apply the multinomial logit
model [@McFadden1974], the expected consumer surplus enjoyed by this individual 
can be shown to be 
\begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation}
where parks are differentiated from each other by their relative measurable
*utilities* $V_{ij}$. $C$ is an unknown constant, but the difference in consumer
surplus between two points $i$ and $k$ can be quantitatively compared as 
$A_i - A_k$. In principle, $V$ may include any measurable attributes of either
the choice maker or the park, and is typically represented as a linear-in-parameters
function of destination attributes $X_j$ and the travel cost $d_{ij}$,
\begin{equation}\label{eq:utility}
V_{ij} = X_j\beta + \beta_d * d_{ij}
\end{equation}
the census tract $i$ in miles. The coefficients $\vect{\beta}$ can be
are frequently estimated from household surveys, though in the absence of a survey we
may assert reasonable values.

Note that the gravity formulation is itself a special case of the utility-based
specification where $\exp(X_j\beta + \beta_d d_{ij}) = X_j f(d_{ij})$ [@Daly1982].
The distinction between gravity and utility-based specification is meaningful, 
however. Primarily, it becomes possible to construct accessibility statistics based
on revealed behavioral preferences rather than calibrated or asserted values.

Both gravity and utility-based specifications hold several advantages relative to
isochrone-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher.
This allows for the fact that some people are more or less sensitive to
distances, and that distance is a continuous, non-binary phenomenon.
It defies reason to assume a person living 1.1 miles
from a park has functionally different accessibility than someone living at 0.9 miles.
Second, the random utility formulation
allows the researcher to include any attribute of the park as part of its utility
specification; in this case, we consider the size of the park weighted against its
distance as an element of relative accessibility. This suggests that
not all parks are equal, and that a large park such as Central Park in Manhattan
may provide health and activity benefits over a much larger area than a smaller
community square.

In spite of its flexibility and basis in choice theory, accessibility measures
based in choice maker utility have not received much application in the accessibility
literature compared with distance-based or even gravity-based measures.
@Vale2016 also present a descriptive classification of
active accessibility mechanisms, though these authors curiously dismiss
utility-based metrics for incorporating randomness. The accessibility
formula presented in \ref{eq:utility} is the expectation of a random utility process,
and is not in and of itself random any more than the gravity model.
@Zhang2011 present a metric with a similar functional form, though the
authors arrive at it from a motivation other than destination choice theory.
Log-sums are commonly used, however, in alternatives analyses of transit
infrastructure improvements [@DeJong2007]. A reason for this is likely that
a regional travel demand model is available to the analysts, thus making
calibrated and multimodal logsums readily available [@Geurs2010].

# Empirical Application
In this section, we develop a model with data for New York City, where
we compute a holistic logsum-based accessibility to parks measure and model the
relationship between this measure and physical activity rates, controlling for
spatial effects and socioeconomic factors. We subsequently model the effect
of the accessibility metric on obesity rates, accounting for physical activity and
the other controls. Finally, we model the relationship between a modified logsum
measure considering geolocated Twitter data and physical activity and obesity rates.

@@

In a destination choice framework, the tracts represent the ``origins'' and
``destinations'' are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We also consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these sub-facilities, we created 
variables for each park indicating the presence of sports courts, playgrounds,
and trails.

@@

With this relationship ruled out, it remains a possibility that the 
model residuals are spatially dependent,
\begin{equation}\label{eq:SEM}
 y = X\beta + \beta_a A + u; u = \lambda W u + \varepsilon
\end{equation}
that the outcome is partially dependent on the socioeconomic variables in *neighboring*
tracts,
\begin{equation}\label{eq:SLX}
 y = X\beta + \beta_a A + W X \lambda + \varepsilon
\end{equation}
or a linear combination of both.
\begin{equation}\label{eq:SDEM}
 y = X\beta + \beta_a A + W X \gamma + u; u = \lambda W u + \varepsilon 
\end{equation}
These models are referred as the spatial error model (SEM), the spatial lag of
$X$ model (SLX), and the spatial Durbin error model (SDEM). @LeSage2014 suggest
that a Hausman-style test of the estimates of $\beta$ derived from the OLS 
(Equation \ref{eq:themodel}) and SEM (Equation \ref{eq:SEM}) can identify whether
the estimates are consistent. If the $\beta$ estimates are dissimilar, both 
specifications are inappropriate as they contain untreated omitted variables.
Similarly, if the SLX and SDEM estimates are dissimilar it is an indication of 
missing variables in that specification. On the other hand if the estimates of
both pairs of models are similar and the residual correlation parameter $\lambda$
is non-zero, the model accounting for spatially dependent residuals will produce 
proper estimates of the model standard errors.

@@

```{r results}
pa_models <- read_rds(here("data/pa_models.rds"))
obesity_models <- read_rds(here("data/obesity_models.rds"))
```

>>>>

Metropolitan-level analyses such as those described above do not capture the
within-city variation in accessibility to parks that exist in many cities. Some
cities with large amounts of total greenspace may nevertheless unequally
distribute this space throughout the city, leading to areas with poor access.
Conversely, a city with a smaller overall proportion of greenspace may give all
of its citizens better access to sufficient greenspace to meet their needs or
wants for physical activity. Considering access at the neighborhood level within
a single city may also eliminate some regional or cultural fixed effects
affecting metropolitan-level analyses.

@@

Loosely defined, the *accessibility* of an arbitrary place describes the ease
with which people can accomplish activities there. Accessibility is an abstract
concept with tempting quantifiability [@Handy1997], perhaps explaining the proliferation of
calculation mechanisms. @Dong2006 present a helpful mathematical heirarchy of
common mechanisms, which we follow here. Consider a person residing at
point $i$ in a city with parks $j \in 1 \ldots J$. An analyst might consider
point $i$ as "having access" to a park if the distance to the park $d_ij$  within an
"isochrone," or less than some threshold $D$,

\begin{equation}\label{eq:isochrone}
A_i = \begin{cases} 
      1 & d_{ij}\leq D\\
      0 & d_{ij} > D
   \end{cases}
\end{equation}

Variations of this isochrone-based framework are easily derived and relatively
common. 
@Dias2019 considered road safety perception as an extenuating factor in the
effective distance to a park.
ParkScore [@parkscore2019] is a sophisticated application of this approach where $d_{ij}$ is a
carefully calculated 10-minute network-bound walk, with a demographic analysis of
areas within and without this threshold. In other applications $D$ might be more
variably defined, such as the presence or percent of green space within a
political or statistical boundary [@Mitchell2008].
The indicator function could also be modified into the total number of parks or
percent of park space within radius $D$ of point $i$ [@Kascynski2014].
In a study of neighborhoods in New York
City, @Stark2014 found that a higher proportion of park space in a neighborhood
was associated with a lower BMI for its residents, and lower park cleanliness
scores were associated with a higher BMI, suggesting that increasing the supply
of clean, well-maintained parks could positively impact adult BMI.

In spite of the flexibility of adapted isochrone techniques, the arbritrary
definitions necessarily imposed by researchers can inhibit holistic analysis,
may have been inititially motivated by computational limitations [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of a point is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_ij)$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function. 
@Giles-Corti2005 compare gravity-based accessibility scores with an isochrone
specification and show that the former are more predictive of walking behavior;
that is, the attributes of a park matter influence walking more than merely the
distance to it. @Zhang2011 present a population-weighted gravity-based
accessibility to parks metric with a national rather than metropolitan scope, 
though they do not examine its correlation with health outcomes.

A third accessibility mechanism is a utility-based specification, termed as such
for being derived from random utility choice theory. Consider that an individual
living at point $i$ is choosing a park for a recreation activity. If we apply
the multinomial logit model [@McFadden1974], the expected consumer surplus
enjoyed by this individual can be shown to be
\begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation}
where parks are differentiated from each other by their relative measurable
*utilities* $V_{ij}$. $C$ is an unknown constant, but the difference in consumer
surplus between two points $i$ and $k$ can be quantitatively compared as 
$A_i - A_k$ [@Bruce1977]. In principle, $V$ may include any measurable attributes of either
the choice maker or the park, and is typically represented as a linear-in-parameters
function of destination attributes $X_j$ and the travel cost $d_{ij}$,
\begin{equation}\label{eq:utility}
V_{ij} = X_j\beta + \beta_d * d_{ij}
\end{equation}
the census tract $i$ in miles. The coefficients $\vect{\beta}$ can be
are frequently estimated from household surveys, though in the absence of a survey we
may assert reasonable values.

Note that the gravity formulation is itself a special case of the utility-based
specification where $\exp(X_j\beta + \beta_d d_{ij}) = X_j f(d_{ij})$ [@Daly1982].
The distinction between gravity and utility-based specification is meaningful, 
however. Primarily, it becomes possible to construct accessibility statistics based
on revealed behavioral preferences rather than calibrated or asserted values [@Handy1997].

Both gravity and utility-based specifications hold several advantages relative
to isochrone-based accessibility metrics more commonly found in the literature.
First, all individuals are defined as having some access to all parks, rather
than an arbitrary cutoff asserted by the researcher. This allows for the fact
that some people are more or less sensitive to distances, and that distance is a
continuous, non-binary phenomenon. It defies reason to assume a person living
1.1 miles from a park has functionally different accessibility than someone
living at 0.9 miles. Second, the random utility formulation allows the
researcher to include any attribute of the park as part of its utility
specification. This suggests that not all parks are equal, and that a large park
such as Central Park in Manhattan may provide health and activity benefits over
a much larger area than a smaller community square.

In spite of its flexibility and basis in choice theory, utility-based
accessibility measures have not received much application in the accessibility
literature compared with distance-based or even gravity-based measures.
@Vale2016 present a descriptive classification of active accessibility techniques, and
explicitly dismiss utility-based metrics for incorporating randomness. The
accessibility formula presented in \ref{eq:utility} is the expectation of a
random utility process, and is not in and of itself random any more than the
gravity model is random.
Utility-based accessibility metrics are commonly used, however, in alternatives
analyses of transit infrastructure improvements [@DeJong2007]. A reason for this
is likely that a regional travel demand model is available to the analysts, thus
making calibrated and multimodal logsums readily available [@Geurs2010].

# Empirical Application
In this section, we develop a model with data for New York City, where
we compute a set of utility-based accessibility to parks scores for each tract
and model the relationship between this measure and physical activity rates,
controlling for spatial effects and socioeconomic factors. We subsequently model
the effect of the accessibility metric on obesity rates, accounting for physical
activity and the other controls.

@@

In a destination choice framework, the tracts represent the "origins" and
"destinations" are parks and green spaces in New York City.
We retrieved a polygons layer of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance [@nycparks].
Upon inspection, we removed several facilities that do not qualify as
publicly accessible green space, such as Yankee Stadium, Citi Field, and their
surrounding parking lots. We also removed parks of less than half an acre in
size, as these appear to be predominantly planted medians rather than legitimate
public green space. We also consolidated individual geographic polygons
comprising a single facility --- as is the case in Flushing Meadows --- and
eliminated sub-facilities such as tennis courts or baseball fields included within
larger park facilities. Instead of these sub-facilities, we created 
variables for each park indicating the presence of sports courts, playgrounds,
and trails.

@@

With this relationship ruled out, it remains a possibility that the 
model residuals are spatially dependent,
\begin{equation}\label{eq:SEM}
 y = X\beta + \beta_a A + u; u = \lambda W u + \varepsilon
\end{equation}
that the outcome is partially dependent on the socioeconomic variables in *neighboring*
tracts,
\begin{equation}\label{eq:SLX}
 y = X\beta + \beta_a A + W X \lambda + \varepsilon
\end{equation}
or a linear combination of both.
\begin{equation}\label{eq:SDEM}
 y = X\beta + \beta_a A + W X \gamma + u; u = \lambda W u + \varepsilon 
\end{equation}
These models are referred as the spatial error model (SEM), the spatial lag of
$X$ model (SLX), and the spatial Durbin error model (SDEM). @Pace2008 suggest
that a Hausman-style test of the estimates of $\beta$ derived from the OLS 
(Equation \ref{eq:themodel}) and SEM (Equation \ref{eq:SEM}) can identify whether
the estimates are consistent. If the $\beta$ estimates are dissimilar, both 
specifications are inappropriate as they contain untreated omitted variables.
Similarly, if the SLX and SDEM estimates are dissimilar it is an indication of 
missing variables in that specification. On the other hand if the estimates of
both pairs of models are similar and the residual correlation parameter $\lambda$
is non-zero, the model accounting for spatially dependent residuals will produce 
proper estimates of the model standard errors.

@@

```{r results}
slx_vs_sdem <- read_rds(here("data/slx_vs_sdem.rds"))
pa_models <- read_rds(here("data/pa_models.rds"))
obesity_models <- read_rds(here("data/obesity_models.rds"))
```
